---
title: "Untitled"
author: "XINLU ZHOU"
date: 10/25/2018
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## K-means clustering

Our first example with **kmeans()** function.

```{r}
# example plot to see how Rmarkdown works
plot(1:10, typ="l")
```

Back to kmeans...

```{r}
# Generate some example data for clustering
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))

plot(x)
```

 Use the kmeans() function setting k to 2 and nstart=20
 Inspect/print the results

```{r}
k <- kmeans(x, centers = 2, nstart = 20)

```
Q. How many points are in each cluster?

```{r}
k$size
```
Q. What ‘component’ of your result object details
      - cluster size?
      `K$size`
      - cluster assignment/membership?
      - cluster center?
```{r}
k$cluster
```

```{r}
table(k$cluster)
```

```{r}
k$centers
```

 Plot x colored by the kmeans cluster assignment and
      add cluster centers as blue points
```{r}
palette(c("blue", "red"))
plot(x, col=k$cluster)
points(k$centers, col = "blue", pch=20, cex=3)
```
Q. Repeat for k=3, which has the lower tot.withinss?
 

```{r}
k2 <- kmeans(x, centers = 2, nstart = 20)
k3 <- kmeans(x, centers = 3, nstart = 20)
k4 <- kmeans(x, centers = 4, nstart = 20)
k5 <- kmeans(x, centers = 5, nstart = 20)

k2$tot.withiness
k3$tot.withinss
k4$tot.withinss
k5$tot.withinss

plot(c(k2$tot.withinss,k3$tot.withinss,k4$tot.withinss,k5$tot.withinss),typ= "b")
```

# Hierarchical clustering
```{r}
d <- dist(x)
hc<- hclust(d)
plot(hc)

```

```{r}
plot(hc)
abline(h=8,col="red")

#cut the tre to yield cluster membership vector
cutree(hc, h=8)
cutree(hc, k=2)
```

```{r}
# Step 1. Generate some example data for clustering
x <- rbind(
  matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2),   # c1
  matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
  matrix(c(rnorm(50, mean = 1, sd = 0.3),           # c3
           rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
# Step 2. Plot the data without clustering
plot(x)
# Step 3. Generate colors for known clusters
#         (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
palette(c("red","blue","black"))
plot(x, col=col,pch=20)
```
# Q. Use the dist(), hclust(), plot() and cutree() functions to return 2 and 3 clusters
```{r}
d <- dist(x)
hc<- hclust(d)
plot(hc)
```
# use cutree function to return 2 and 3 clusters
```{r}
grp.2 <- cutree(hc, k=2)
grp.3 <- cutree(hc, k=3)

```
# Q. How does this compare to your known 'col' groups?
```{r}
table(grp.3,col)
```

#principal component analysis
#Let's make up some data data
```{r}
## Initialize a blank 100 row by 10 column matrix
mydata <- matrix(nrow=100, ncol=10)
## Lets label the rows gene1, gene2 etc. to gene100
rownames(mydata) <- paste("gene", 1:100, sep="")
## Lets label the first 5 columns wt1, wt2, wt3, wt4 and wt5
##   and the last 5 ko1, ko2 etc. to ko5 (for "knock-out")
colnames(mydata) <- c( paste("wt", 1:5, sep=""),
                       paste("ko", 1:5, sep="") )
## Fill in some fake read counts
for(i in 1:nrow(mydata)) {
wt.values <- rpois(5, lambda=sample(x=10:1000, size=1)) 
ko.values <- rpois(5, lambda=sample(x=10:1000, size=1))
mydata[i,] <- c(wt.values, ko.values)
}
head(mydata)
```

```{r}
head(t(mydata))
```

```{r}
pca <- prcomp(t(mydata), scale=TRUE)
pca
```
Make a PC plot of pc1 vs pc2. for this i used "$x" component of our output
```{r}
plot(pca$x[,1],pca$x[,2] )
```

```{r}
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)
barplot(pca.var.per, main="Scree Plot",
xlab="Principal Component", ylab="Percent Variation")
```

```{r}
## A vector of colors for wt and ko samples
colvec <- colnames(mydata)
colvec[grep("wt", colvec)] <- "red"
colvec[grep("ko", colvec)] <- "blue"
plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
     xlab=paste0("PC1 (", pca.var.per[1], "%)"),
     ylab=paste0("PC2 (", pca.var.per[2], "%)"))
```
   
```{r}
plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
     xlab=paste0("PC1 (", pca.var.per[1], "%)"),
     ylab=paste0("PC2 (", pca.var.per[2], "%)"))
```


#food pca
```{r}
x <- read.csv("UK_foods.csv")
dim(x)
```
```{r}
head(x)
```
```{r}
rownames(x) <- x[,1]
x <- x[,-1]
head(x)
```

```{R}
dim(x)
```
```{r}
knitr::kable(x, caption="The full UK foods data table")
```

```{r}
heatmap(as.matrix(x))
```

pca <- prcomp( t(x))
summary(pca)
